\documentclass[]{article}
\input{../mathdoc}
\input{../mathsym}
\input{../preview}

\author{Book: Thomas Judson 2017, Presenter: Igor Erovenko, Notes by Michael Reed}
\title{Abstract Algebra}
%date{}

\begin{document}
\maketitle

%\begin{abstract}
%\end{abstract}

\section{Proofs}

Let $p$ and $q$ be statements. \say{If $p$ then $q$.} $p$ is the hypothesis and $q$ is the conclusion. \ul{Implication} is $p\implies q$, read \say{$p$ implies $q$.}

\begin{tabular}{|c|c|c|}
	\hline 
	$p$ & $q$ & $p\implies q$ \\ 
	\hline 
	T & T & T \\ 
	\hline 
	T & F & F \\ 
	\hline 
	F & T & T \\ 
	\hline 
	F & F & T \\ 
	\hline 
\end{tabular}

\subsection*{Methods of proof}

\begin{enumerate}
	\item Direct proof. Assume $p$ is true, show $q$ is true. $p\implies \dots \implies \dots \implies\dots \implies q$.
	\begin{example}
		Show that if $n$ is an odd integer, then $n^2$ is odd. $p = $ \say{$n$ is odd} and $q = $ \say{$n^2$ is odd.}
		\begin{proof}
			Suppose $n$ is an odd integer. Then $n = 2k+1$ for some integer $k$. In this case, $n^2 = (2k+1)^2 = 4k^2 + 4k + 1 = 2(2k^2 + 2k) + 1$. Notice that $2k^2+2k$ is an integer. So, $n^2$ is odd.
		\end{proof}
	\end{example}
	\item Reduction to cases. If $p$ is true, then one of $p_1,p_2,\dots,p_n$ must be true. Then show each $p_i \implies q$.
	\begin{example}
		Show that if $n$ is an integer, then $n^2-n$ is even. $p = $ \say{$n$ is an integer} and $q = $ \say{$n^2-n$ is even.}
		\begin{proof}
			Suppose $n$ is an integer. Then $n$ is even or odd.
			\ul{Case 1}: $n$ is even. In this case, $n=2k$ for some integer $k$ and $n^2-n = (2k)^2 - 2k = 4k^2 - 2k = 2(2k^2-k)$ is even. \ul{Case 2}: $n$ is odd. In this case $n = 2k+1$ for some integer $k$ and $n^2-n = (2k+1)^2-(2k+1) = 4k^2 + 4k + 1 - 2k -1 = 4k^2 +2k = 2(2k+k)$ is even.
		\end{proof}
	\end{example}
	\item Proof by contradiction. If $p$ is true, then either $q$ is true or it is not true. Show that $\neg q$ is impossible. Assume: $p$ is true and $q$ is false (not $q$ is true). Show: this leads to a contradiction.
	\begin{example}
		Show that $\sqrt{2}$ is irrational. If $r$ is rational, then $r^2 \neq 2$.
		\begin{proof}
			Assume $r$ is rational and $r^2 = 2$. Then $r = \frac{m}{n}$ for some integers $m$ and $n$, in lowest terms. Then $(\frac{m}{n})^2 = 2$. So, $m^2 = 2n^2$. It follows that $m^2$ is even, and hence $m$ is even. Then $m = 2k$ and we get $(2k)^2 = 4k^2 = 2n^2$ or $2k^2 = n^2$. Therefore $n^2$ is even, and consequently $n$ is even. Both $m$ and $n$ are even, which contradicts our assumption that $\frac{m}{n}$ is in lowest terms. Therefore $r^2 \neq 2$ and $\sqrt{2}$ is irrational.
		\end{proof}
	\end{example}
\end{enumerate}

To disprove a general statement: find one counterexample.

\begin{example}
	Every integer is odd. False, $4$ is not odd.
\end{example}

Equivalences: $p$ if and only if $q$. $p\Longleftrightarrow q$, needs proofs for both $p\implies q$ and $q\implies p$.

\section{Sets}

A \ul{set} is a well-defined collection of objects. Requires a criterion for determining set membership. $a\in A$ means $a$ is a \ul{member} or an \ul{element} of a set $A$.

$$X = \{ x_1,x_2,\dots,x_n\}, \, X = \{ x: x \text{ satisfies } P\} = \{x \mid x \text{ satisfies } P\}.$$

$\mb N$ natural numbers $\{1,2,3,4,\dots\}$.
$\mb Z$ integers $\{\dots,-3,-2,-1,0,1,2,3,\dots\}$.
$\mb Q$ rational numbers $\{\frac{m}{n} \mid m,n\in \mb Z, n\neq 0\}/\sim$.
$\mb R$ real numbers.
$\mb C$ complex numbers $\{a+bi \mid a,b\in\mb R\}$.
$\mb H$ Hamiltonian quaternions $\{a+bi+cj+dk \mid a,b,c,d\in\mb R\}$.

\begin{definition}
	[Subset] $A\subset B$ or $A\subseteq B$, $a\in A \implies a\in B$.
\end{definition}
\begin{definition}
	[Equality of sets]
	$A = B \Longleftrightarrow A \subseteq B$ and $B\subseteq A$.
\end{definition}

\begin{definition}
	[Proper subset] $A\subsetneq B \Longleftrightarrow A\subseteq B$ and $A\neq B$.
\end{definition}
\begin{definition}
	[Empty set] $\emptyset \subseteq A$ for any set $A$.
\end{definition}

\begin{definition}
	[Union] $A\cup B = \{ x\mid x\in A \text{ or } x\in B\}$.
\end{definition}
\begin{definition}
	[Intersection] $A\cap B = \{x \mid A \text{ and } x\in B\}$.
\end{definition}
\begin{definition}
	[Difference] $A-B = A\backslash B = \{x \mid x\in A \text{ and } x\notin B\}$.
\end{definition}

\begin{proposition}
	Let $A,B,$ and $C$ be sets.
	\begin{enumerate}
		\item $A\cup A = A$, $A\cap A = A$, $A\backslash A = \emptyset $.
		\item $A\cup \emptyset = A$, $A\cap \emptyset = \emptyset$.
		\item $A\cup (B\cup C) = (A\cup B)\cup C$.
		$A\cap (B\cap C) = (A\cap B) \cap C$.
		\item $A\cup B = B\cup A$. $A\cap B = B\cap A$.
		\item $A\cup (B\cap C) = (A\cup B) \cap (A\cup C)$.
		\begin{proof}
			Let $x\in A\cup (B\cap C)$. Then $x\in A$ or $x\in B\cap C$. \ul{Case 1}: $x\in A$. In this case, $x\in A\cup B$ and $x\in A\cup C$, so $x\in (A\cup B)\cap (A\cup C)$. \ul{Case 2}: $x\in B\cap C$. In this case, $x \in B$ and $x\in C$. So, $x\in A\cup B$ and $x\in A\cup C$. Hence $x \in (A\cup B)\cap (A\cup C)$. Therefore $A\cup (B\cap C) \subseteq (A\cup B)\cap (A\cup C)$.
		\end{proof}
		\item $A\cap (B\cup C) = (A\cap B) \cup (A\cap C)$.
	\end{enumerate}
\end{proposition}

\subsection*{Cartesian products and mappings}

Cartesian product of $A$ and $B$ is $A\times B = \{(a,b)\mid a\in A \text{ and } b\in B\}$.
In general, $A_1\times A_2\times\dots\times A_n = \{(a_1,a_2,\dots,a_n)\mid a_i\in A_i \text{ for } 1\leq i \leq n\}$.
\ul{Notation}: $A\times A\times \dots\times A = A^n$.

\begin{example}
	$\mb R^2, \mb R^3$.
\end{example}

A \ul{relation} from $A$ to $B$ is a subset of $A\times B$.
A \ul{mapping} (or a \ul{function}) $f$ from $A$ to $B$ is a relation from $A$ to $B$ such that for every $a\in A$ there exists a unique $b\in B$ for which $(a,b)\in f$. \ul{Notation}: $(a,b) \in f \ra f(a) = b$ or $f:a\mapsto b$ and on the level of sets $f:A\ra B$. $A$ is the domain of $f$, $B$ is the codomain, $f(A) = \{f(a)\mid a\in A\}\subseteq B$ is the range of $f$.

\begin{definition}
	A mapping $f:A\ra B$ is
	\begin{enumerate}
		\item \ul{injective} (or one-to-one) if $f(a_1) = f(a_2) \implies a_1 = a_2$.
		\item \ul{surjective} (or onto) if $f(A) = B$, or for every $b\in B$ there exists $a\in A$ such that $f(a) = b$.
		\item \ul{bijective} (or one-to-one correspondence) if it's injective and surjective.
	\end{enumerate}
\end{definition}
\begin{example}
	$f:\mb R\ra \mb R$, $f(x) = e^x$. $e^{x_1} = e^{x_2} \overset{?}{\implies} x_1 = x_2$ then $\ln(e^{x_1}) = \ln(e^{x_2})$, so injective. $f(\mb R) = \mb R^+ = (0,\infty) \neq \mb R$, not surjective and not bijective.
	$f:\mb R \ra (0,\infty)$ is bijective.
\end{example}
\begin{example}
	$f:\mb Z\ra\mb Z$, $f(n) = n^2$. $f(-1) = 1 = f(1)$, not injective. $2\notin f(\mb Z)$, not surjective and not bijective.
\end{example}
\begin{example}
	$f:\mb{R}\ra \mb R$, $f(x) = 2x+1$. $2x_1 + 1 = 2x_2 + 1 \implies x_1 = x_2$, injective. Given any $y\in \mb R$ find $x\in \mb R$ such that $2x+1 = y$: $x = (y-1)/2$, surjective and bijective.
\end{example}

A \ul{composition} of mappings $f:A\ra B$ and $g:B\ra C$ is a mapping $g\circ f: A\ra C$ defined by $(g\circ f)(a) = g(f(a))$. $A \overset{f}{\lra} B \overset{g}{\lra} C$ or $a \rightsquigarrow f(a) \rightsquigarrow g(f(a))$.
\begin{theorem}
	Let $f:A\ra B$, $g:B\ra C$, and $h:C\ra D$ be mappings.
	\begin{enumerate}
		\item The composition of mappings is associative: $(h\circ g) \circ f = h\circ(g\circ f)$.
		\begin{proof}
			Let $a\in A$. Then $((h\circ g)\circ f)(a) = (h\circ g)(f(a)) = h(g(f(a))) \equiv h((g\circ f)(a)) = (h\circ(g\circ f))(a)$.
		\end{proof}
		\item If $f$ and $g$ are injective, then so is $g\circ f$.
		\begin{proof}
			$(g\circ f)(a_1) = (g\circ f)(a_2) \implies g(f(a_1)) = g(f(a_2)) \implies f(a_1) = f(a_2) \implies a_1 = a_2$.
		\end{proof}
		\item If $f$ and $g$ are surjective, then so is $g\circ f$.
		\begin{proof}
			$A\os{f}{\lra} B \os{g}{\lra} C$ or $A\os{g\circ f}{\lra} C$. $g(b) = c$, $f(a) = b$.
		\end{proof}
		\item If $f$ and $g$ are bijective, then so is $g\circ f$.
	\end{enumerate}
\end{theorem}

\begin{definition}
	The \ul{identity} mapping on a set $A$ is $\text{id}_A: A\ra A$, $\text{id}_A(a) = a$.
	An \ul{inverse} mapping of a mapping $f:A\ra B$ is a mapping $g:B\ra A$ such that $g\circ f = \text{id}_A \ra g(f(a)) = a$, $f\circ g = \text{id}_B \ra f(g(b)) = b$.
	A mapping is \ul{invertible} if it has an inverse mapping. \ul{Notation}: $f\inv$ inverse of $f$.k
\end{definition}

\begin{theorem}
	A mapping is invertible $\Leftrightarrow$ it is bijective.
\end{theorem}

\subsection*{Equivalence relations}

\begin{definition}
	An \ul{equivalence relation} on a set $X$ is a relation $R\subset X\times X$ which is
	\begin{enumerate}
		\item \ul{Reflexive}: $(x,x)\in R$ for all $x\in X$
		\item \ul{Symmetric}: $(x,y)\in R \implies (y,x)\in R$
		\item \ul{Transitive}: $(x,y)\in R,(y,z)\in R\implies (x,z)\in R$
	\end{enumerate}
	\ul{Notation}: $x\sim y$ for $(x,y)\in R$.
\end{definition}
\begin{example}
	$X=\mb R$, $a\sim b$ if $a<b$. Reflexive: $0\nless 0$. Symmetric: $0<1$ but $1\nless 0$. Transitive: $a<b$ and $b<c\implies a<c$. However, for $a\neq b$ either $a<b$ or $b<a$ (antisymmetric).
\end{example}
\begin{example}
	$X = \mb R$, $a\sim b$ if $ab\geq 0$. \ul{Reflexive}: $a\sim a \Leftrightarrow a^2 \geq0$. \ul{Symmetric}: $a\sim b \implies b\sim a$ or $ab\geq 0\implies ba\geq 0$. \ul{Transitive}: $a\sim b,b\sim c\implies a\sim c$ or $ab\geq 0,bc\geq 0 \implies ac\geq 0$. Counterexample: $a=-1,b=0,c=1$.
\end{example}
\begin{example}
	$X$ any set, $x\sim y$ if $x=y$. \ul{Reflexive}: $x\sim x \Leftrightarrow x=x$. \ul{Symmetric}: $x\sim y\implies y\sim x$ or $x=y\implies y=x$. \ul{Transitive}: $x\sim y,y\sim z\implies x\sim z$ or $x=y,y=z\implies x=z$.
\end{example}
\begin{example}
	$X = $ all differentiable functions $\mb R\ra\mb R$. $f\sim g$ if $f' = g'$. \ul{Reflexive}: $f\sim f \Leftrightarrow f' = f'$. \ul{Symmetric}: $f\sim g \implies g\sim f$ or $f' = g' \implies g' = f'$. \ul{Transitive}: $f\sim g,g\sim h\implies f\sim h$ or $f' = g',g'=h'\implies f'=h'$.
\end{example}
\begin{example}
	$X = \mb R^2$, $(x_1,y_1)\sim (x_2,y_2)$ if $x_1^2 + y_1^2 = x_2^2 + y_2^2$.
\end{example}
\begin{example}
	$X = M_n(\mb R)$ all $n\times n$ matrices. $A\sim B$ if $\exists$ a nonsingular P such that $B = PAP\inv$. \ul{Reflexive}: $A = IAI\inv$ let $P = I$. \ul{Symmetric}: $A\sim B \implies B\sim A \lra A = *B*\inv$, then $B = PAP\inv \implies P\inv B = AP\inv \implies P\inv BP = A$, so let $* = P\inv$. \ul{Transitive}: $A\sim B,B\sim C\implies A\sim C$. $B = PAP\inv, C = SBS\inv \implies C = *A*\inv$. $C = SBS\inv = S(PAP\inv)S\inv = (SP)A(P\inv S\inv) = (SP)A(SP)\inv$, so $*= SP$.
\end{example}

\newpage

\begin{definition}
	Let $\sim$ be an equivalence relation on a set $X$, and let $x\in X$. Then the set $$[x] = \{y\in x \mid y\sim x\}$$ is the \ul{equivalence class} of $x$.
\end{definition}
\begin{example}
	$X$ is any set. $x\sim y$ if $x=y$. Then $[x] = \{y\in X\mid y=x\} = \{x\}$.
\end{example}
\begin{example}
	$X$ is the set of all differentiable functions from $\mb R\ra\mb R$. $f\sim g$ if $f' = g'$. $[f] = \{g\in X\mid g' = f'\} = \{f + c\mid c\in\mb R\}$.
\end{example}
\begin{example}
	$X = \mb R^2$. $(x_1,y_1)\sim (x_2,y_2)$ if $x_1^2 + y_1^2 = x_2^2 + y_2^2$. The equivalence class is a circle.
\end{example}

\begin{definition}
	A \ul{partition} of a set $X$ is a collection of disjoint nonempty subsets of $X$ which cover $X$.
	$$\mc P = \{X_\alpha\}_{\alpha\in \Lambda}, \quad X\alpha\subseteq X\, \forall \alpha,$$
	$X_\alpha \cap X_\beta = \emptyset \text{ if } \alpha \neq \beta$, disjoint, and $X = \bigcup_{\alpha\in\Lambda} X_\alpha$ cover $X$.
\end{definition}
\begin{theorem}
	Let $\sim $ be an equivalence relation on a set $X$. Then the equivalence classes of $\sim$ partition $X$. Conversely, given a partition $\mc P = \{X_\alpha\}_{\alpha\in\Lambda}$ of $X$, there exists an equivalence relation $\sim$ on $X$ such that $\{X_\alpha\}$ are the equivalence classes of $\sim$.
\end{theorem}
\begin{proof}
	Let $\sim$ be an equivalence relation on $X$. Then for every $x\in X$, $x\sim x$, so $x\in [x]$. Hence, $[x]\neq \emptyset$ for all $x\in X$. Also $\bigcup_{x\in X}[x] = X$. By definition $\bigcup_{x\in X}[x] \subseteq X$ and  every $x$ is in one of equivalence classes so $\bigcup_{x\in X}[x] \supseteq X$. It's left to prove that if $x,y\in X$ then $[x]\cap [y] = \emptyset$ or $[x]=[y]$.
	\ul{Case 1}: $[x]\cap[y] = \emptyset$ done. \ul{Case 2}: $[x]\cap[y]\neq \emptyset$ (WTS: $[x]=[y]$). There is $z\in[x]\cap[y]$. Then $z\in[x]$ and $z\in[y]$, so $z\sim x$ and $z\sim y$. By symmetric property $x\sim z$ and by transitive property $x\sim y$. Take any $a\in [x]$. Then $a\sim x$ and also $x\sim y$, so $a\sim y$ and thus $a\in [y]$. So $[x]\subseteq [y]$. If $x\sim y$ then $y\sim x \implies [y]\subseteq [x]$. Thus $[x]=[y]$.
	
	Conversely, let $\mc P = \{X_\alpha\}_{\alpha\in\Lambda}$ be a partition of $X$. Define a relation $\sim$ on $X$ by $x\sim y$ if $x$ and $y$ belong to the same $X_\alpha$ for some $\alpha\in\Lambda$. Show:
	\begin{enumerate}
		\item $\sim$ is an equivalence relation on $X$,
		\item the equivalence classes of $\sim$ are $X_\alpha$'s
	\end{enumerate}
\end{proof}

\section*{Mathematical induction}

$$P(1),P(2),P(3),\dots,P(n),P(n+1),\dots$$

Mathematical induction: prove that $P(n)$ is true for all $n\geq 1$.

\begin{definition}
	[Well-Ordering Principle] Every nonempty subset of $\mb N$ contains a smallest element.
\end{definition}
\begin{theorem}
	[First principle of mathematical induction] Suppose $P(n)$ is a statement involving a positive integer $n$, and
	\begin{enumerate}
		\item \textit{Base}: $P(n_0)$ is true for some $n_0\in\mb N$
		\item \textit{Induction hypothesis}: $P(k)$ implies $P(k+1)$ for all $k\geq n_0$.
	\end{enumerate}
	$P(n)$ is true for all $n\geq n_0$.
\end{theorem}
\begin{remark}
	This is nothing but a \say{domino} effect. $P(k) = $ \say{domino $k$ falls}.
\end{remark}
\begin{proof}
	Assume on the contrary that not all $P(n)$ are true for $n\geq n_0$. Then $S = \{n\geq n_0\mid P(n) \text{ is false}\}\subseteq \mb N$ is a nonempty subset of $\mb N$. By the Well-Ordering Principle, the set $S$ contains a least element $m$. Then $P(m)$ is false and $m\neq n_0$ by assumption (1). This means that $m>n_0 \implies m-1\geq n_0$. So $P(m-1)$ is true since $m-1\notin S$. By assumption (2), if $P(m-1)$ is true, then $P(m)$ is true, which is a contradiction.
\end{proof}
\begin{example}
	$1+2+\dots+n = \frac{n(n+1)}{2}$ for all $n\geq 1$. $P(n) = $ \say{$1+2+\dots+n = \frac{n(n+1)}{2}$.}
	\ul{Base}: $P(1)$ is true since $1 = \frac{1\cdot(1+1)}{2}$.
	\ul{Inductive step}: Assume $P(k)$ is true for some fixed arbitrary $k \geq 1$. Want to show: $P(k+1)$ is true. Have: $1+2+\cdots+k = \frac{k(k+1)}{2}$. Want: $1+2+\dots+k + (k+1) = \frac{(k+1)(k+2)}{2}$.
	\begin{align*}
		1+2+\dots + k + (k+1) &= \frac{k(k+1)}{2} + (k+1) = \frac{k(k+1)}{2} + \frac{2(k+1)}{2} \\
		&= \frac{k(k+1) + 2(k+1)}{2} = \frac{(k+1)(k+2)}{2}
	\end{align*}
	Therefore, $P(n)$ is true for all $n\geq 1$.
\end{example}
\begin{example}
	All cars are the same color. $P(n) = $ \say{in any collection of $n$ cars, all cars are the same color.} Will prove by induction that $P(n)$ is true for all $n\geq 1$. \ul{Base}: $P(1)$ is true. \ul{Inductive step}: $P(k)$ is true for some $k\geq 1$. WTS: $P(k+1)$ is true. Cars $c_1,\dots,c_k$ are same color. However $P(1)\nrightarrow P(2)$.
\end{example}

\begin{example}
	The UNCG poker club plays with \$5 and \$8 chips. What is the largest bet that cannot be made using these num? Largest bet that cannot be made is \$27.
	
	We need to show
	\begin{enumerate}
		\item A bet of \$27 cannot be made.
		\item Any bet $\geq \$28$ can be made.
	\end{enumerate}
	A bet of $\$n$ can be made if $n=5a+8b$, where $a$ and $b$ are nonnegative integers. Take this modulo 5, then $2\equiv 3b\pmod5$. Let $b=0$, then $2\not\equiv 0\pmod5$. Let $b=1$, then $2\not\equiv 3\pmod 5$. Let $b=2$, then $2\not\equiv 6\pmod5$. Let $b = 3$, then $2\not\equiv 9\pmod5$. Let $b=4$, then $2\equiv 12\pmod 12$. So $b\geq 4$ and $27 = 5a + 8b\geq 5a+32$ implies $5a\leq -5\implies a\leq -1$. $P(n) = $ \say{A bet of $\$n$ can be made} = \say{$n = 5a+8b$ for some nonnegative integers $a$ and $b$.} WTS: $P(n)$ is true for all $n\geq 28$.
	\ul{Base}: $P(28)$ is true since $28 = 5\cdot 4 + 8\cdot 1$.
	\ul{Inductive step}: Assume $P(k)$ is true for some arbitrary fixed $k\geq 28$. WTS: $P(k+1)$ is true. Assume: $k=5a+8b$. Want: $k+1 = 5\tilde a + 8\tilde b$. If $k = 5a+8b$ then $k+1 = 5a+8b+1$. Then $1 = 5\cdot(-3) + 8\cdot(2) = 5\cdot(5) + 8\cdot(-3)$. Then $5a+8b +1 = 5\cdot(a-3) + 8\cdot(b+2) = 5\cdot(a+5) + 8\cdot(b-3)$.
	\ul{Claim}: $a \geq 3$ or $b\geq 3$. Assume on the contrary $a<3$ and $b<3$. Then $k = 5a+8b \leq 5\cdot 2 + 8\cdot 2 = 26$. This is a contradiction.
	\ul{Case 1}: Let $a\geq 3$, then $k+1 = 5\cdot(a-3) + 8\cdot(b+2)$ is a valid bet.
	\ul{Case 2}: Let $b\geq 3$, then $k+1 = 5\cdot (a+5) + 8\cdot(b-3)$ is a valid bet.
\end{example}

\begin{theorem}
	[Second form of mathematical induction] Suppose $P(n)$ is a statement involving a positive integer $n$, and
	\begin{enumerate}
		\item $P(n)$ is true for some $n_0\in\mb N$.
		\item $P(n_0),P(n_0+1),\dots,P(k)\implies P(k+1)$ for all $k\geq n_0$.
	\end{enumerate}
	Then $P(n)$ is true for all $n\geq n_0$.
\end{theorem}
\begin{proof}
	Let $Q(n) = $ \say{$P(n_0),P(n_0+1),\dots,P(n)$ are true.} Apply the first principle of mathematical induction to the statements $Q(n)$. Let $Q(n_0) = $ \say{$P(n_0)$ is true}, then $Q(k) \implies Q(k+1)$.
\end{proof}
\begin{example}
	Every natural number $\geq 2$ is a product of primes.
	
	\ul{Base}: Let $n=2$, which is a prime.
	\ul{Inductive step}: Assume that all $2,3,\dots,k$ are products of primes. Look at $k+1$. \ul{Case 1}: $k+1$ is prime, then done. \ul{Case 2}: $k+1$ is composite, then $k+1 = uv$ with $u,v\geq 2$. Then $2\leq u,v\leq k$. Let $u = p_1p_2\cdots p_s$ and $v = q_1q_2\cdots q_r$, where $p_i,q_j$ are primes. Then $k+1 = p_1p_2\cdots p_sq_1q_2\cdots q_r$. m
\end{example}

\subsection{The division algorithm}
\begin{theorem}
	[The division algorithm] Let $a$ and $b>0$ be integers. Then there exist unique integers $q$ and $r$ such that $a = qb +r$, where $0\leq r < b$.
\end{theorem}
From now on: $a,b\in\mb Z$. If $a =kb$ for some $k\in\mb Z$ we say $b$ divides $a$ or $b\mid a$. An integer $d$ is a \ul{common divisor} of $a$ and $b$ if $d\mid a$ and $d\mid b$.
\begin{definition}
	The \ul{greatest common divisor} of integers $a$ and $b$ is a \ul{positive} integer $d$ such that
	\begin{enumerate}
		\item $d$ is a common divisor of $a$ and $b$
		\item any common divisor $d'$ of $a$ and $b$ divides $d$.
	\end{enumerate}
	\ul{Notation}: $d = \gcd(a,b) = (a,b)$
\end{definition}
$a$ and $b$ are \ul{relatively prime} if $\gcd(a,b) = 1$.
\begin{theorem}
	Let $a$ and $b$ be two integers ($\neq 0$).
	\begin{enumerate}
		\item $\gcd(a,b)$ is unique
		\item $\exists r,s\in\mb Z$ such that $\gcd(a,b) = ar+bs$.
	\end{enumerate}
	In particular, $\gcd$ exists.
\end{theorem}
\begin{proof}
	\begin{enumerate}
		\item Assume $d_1 = \gcd(a,b)$ and $d_2 = \gcd(a,b)$. WTS: $d_1 = d_2$.
		\begin{enumerate}
			\item[(a)] Assume $d_1$ is $\gcd$ and $d_2$ is common divisor, so $d_2\mid d_1 \implies d_2 \leq d_1$.
			\item[(b)] Assume $d_2$ is $\gcd$ and $d_1$ is common divisor, so $d_1\mid d_2 \implies d_1 \leq d_2$.
		\end{enumerate}
		So $d_1 = d_2$.
		
		\item $S = \{ am + bn \mid m,n\in\mb Z;\ am+bn>0\}\subseteq \mb N$. So $S$ is a nonempty subset of $\mb N$. Let $d$ be the least element of $S$ (Well-Ordering Principle). \ul{Claim}: $d = \gcd(a,b)$. Say $d = ar+bs$, for some $r,s\in\mb Z$. WTS: $d\mid a$ and $d\mid b$. By the division algorithm, $a = dq + \tilde{r}$, $0\leq \tilde{r}<d$. Solve for $\tilde{r} = a-dq = a-(ar+bs)q = a - arq - bsq = a(1-rq)+b(-sq) \geq 0$. Assum $\tilde r >0$. Then $\tilde r \in S$, but $\tilde r<d$, contradiction $\implies \tilde r = 0 \implies d\mid a$. For $d\mid b \ra$ similar to $d\mid a$.
		
		WTS: any common divisor $d'$ of $a$ and $b$ divides $d$. So $a = d' u$ and $b = d' v$ for some $u,v\in\mb Z$. Then $d = ar+bs = (d'u)r + (d'v)s = d'(ur + vs)$, where $ur+rs\in\mb Z$. So $d'\mid d$.
	\end{enumerate}
\end{proof}

\subsection{Euclidean algorithm}
Given: $a$ and $b$, $a>b$.
Find: $d = \gcd(a,b)$ and $d = ar+bs$.

\begin{align*}
	a &= bq_1 + r_1 \\
	b &= r_1q_2 + r_2 \\
	r_1 &= r_2q_3 + r_3 \\
	&\ \,\vdots \\
	r_{n-2} &= r_{n-1}q_n + r_n \\
	r_{n-1} &= r_n q_n
\end{align*}
\ul{Claim}: $r_n = \gcd(a,b)$.
\begin{enumerate}
	\item WTS: $r_n\mid a$ and $r_n\mid b$. Last equation tells $r_n\mid r_{n-1} \implies r_n\mid r_{n-2} \implies \dots \implies r_n\mid b \implies r_n\mid a$.
	\item Assume $d'\mid a$ and $d'\mid b$. First equation tells $d'\mid r_1 \implies d'\mid r_2 \implies d'\mid r_3 \implies \dots \implies d'\mid r_n$.
\end{enumerate}
It stops: (assume $a,b>0$), so $b>r_1 > r_2 > \dots > r_n$.

\begin{example}
	1001 and 357.
	\begin{align*}
		1001 &= 357\cdot 2 + 287\\
		357 &= 287\cdot 1 + 70\\
		287 &= 70\cdot 4 + 7\\
		70 &= 7\cdot 10 
	\end{align*}
	So $7 = \gcd(1001,357)$. Then $7 = 287 - 70\cdot 4 = 287 - (357 - 287)\cdot 4 = 287 \cdot 5 - 357 \cdot 4 = (1001-357\cdot 2)\cdot 5 - 357\cdot 4 = 1001 \cdot 5 + 357 \cdot (-14)$.
\end{example}
\begin{corollary}
	If $a$ and $b$ are relatively prime, then there exist integers $r$ and $s$ such that $ar+bs = 1$.
\end{corollary}
If $d = \gcd(a,b)\implies d = ar+bs$.

An integer $p>1$ is \ul{prime} if the only divisors of $p$ are 1 and $p$. Not prime $\lra$ composite.

\begin{lemma}
	[Euclid] Let $a,b\in\mb Z$ and $p$ a prime number. If $p\mid ab$ then $p\mid a$ or $p\mid b$.
\end{lemma}
\begin{proof}
	\ul{Case 1}: $p\mid a$, done.
	\ul{Case 2}: $p\nmid a$. WTS $p\mid b$. If $p\nmid a$ then $\gcd(a,p) = 1 \implies \exists r,s\in\mb Z$ such that $ar+ps = 1$. Then $b = b\cdot 1 = b (ar+ps) = abr + pbs$. Since $p\mid ab$ and $p\mid pbs$, then $p\mid b$.
\end{proof}
\begin{theorem}
	[Euclid] There are infinitely many primes.
\end{theorem}
\begin{proof}
	Assume there are finitely many primes: $p_1,p_2,\dots,p_n$. Let $p = p_1p_2\dots p_n + 1$. So $p$ is a product of primes. However, $p_i\nmid p,\ 1\leq i\leq n \implies p$ is a prime number, a new prime (contradiction).
\end{proof}
\begin{theorem}
	[Fundamental Theorem of arithmetic] Let $n\geq 2$ be an integer. Then there exist primes $p_1,\dots p_s$ (not necessarily distinct) such that $n = p_1p_2\dots p_s$.
	Moreover, this factorization is unique.
\end{theorem}
\begin{proof}
	Assume $n = p_1 p_2\dots p_s = q_1 q_2\dots q_r$, primes. Notice that $p_1\mid n = q_1 q_2\dots q_r$. Recall Euclid's lemma: $p_1 \mid$ one of $q_i$'s $\implies p_1 = q_1$.
\end{proof}

\subsection*{Modular arithmetic}
Fix $n\in\mb N$. For $a,b\in\mb Z$, $a \equiv b\pmod{n}$ if $a-b$ is divisible by $n$.
Define a relation on $\mb Z$ by $a\sim b$ if $a\equiv b\pmod{n}$.
\ul{Claim}: This is an equivalence relation on $\mb Z$.
\begin{enumerate}
	\item Reflexive: $a\equiv a \pmod{n}$ since $n\mid 0 = a-a$ for all $a\in\mb Z$.
	\item Symmetric: $a\equiv b\pmod{n}\implies b\equiv a\pmod{n}$, since $n\mid(a-b) \lra a-b=nk \lra b-a = n(-k)$.
	\item Transitive: $a\equiv b\pmod{n}$ and $b\equiv c\pmod{n} \implies a\equiv c\pmod{n}$, since $a-b=nk$ and $b-c=nl$ and $a-c = (a-b)+(b-c) = nk+nl = n(k+l)$.
	Let $a = nq_1+r_1$ and $b=nq_2+r_2$, then $a-b = n(q_1-q_2) + (r_1-r_2)$ with $r_1-r_2 = 0$.
\end{enumerate}

\begin{example}
	\ul{$n=3$} Equivalence class:

	$[0] = \{ \dots,-6,-3,0,3,6,9,12,\dots \} \lra $ remainder 0.

	$[1] = \{ \dots,-5,-2,1,4,7,10,13,\dots \} \lra $ remainder 1.

	$[2] = \{ \dots, -4,-1,2,5,8,11,14,\dots \} \lra $ remainder 2.
\end{example}
In general, $\{[0],[1],[2],\dots,[n-1]\} = \mb Z_n$.
Define two operations:
\begin{enumerate}
	\item Addition: $[a]+[b] = [a+b]$.
	\item Multiplication: $[a]\cdot[b] = [ab]$.
	
\end{enumerate}
\begin{example}
	$\mb Z_6 = \{[0],[1],[2],[3],[4],[5]\}$. Take $[3]+[4] = [7]=[1]$, or $[0]+[3] = [0+3] = [3]$. Take $[2]\cdot[4] = [2\cdot 4] = [8] = [2]$, or $[2]\cdot[3] = [2\cdot 3] = [6] = [0]$. Now take $[1]+[2] = [3]$ and $[7]+[14]=[21]$, they are equal.
\end{example}
\begin{lemma}
	Let $a\equiv a'\pmod n$ and $b\equiv b'\pmod n$. Then
	\begin{enumerate}
		\item $a+b \equiv a'+b'\pmod n$,
		\item $ab \equiv a'b'\pmod n$.
	\end{enumerate}
\end{lemma}
\begin{proof}
	$a\equiv a'\pmod n \lra n\mid (a-a')$ and $b\equiv b'\pmod n \lra n\mid(b-b')$.
	\begin{enumerate}
		\item $(a+b) - (a'+b') = (a-a') + (b-b')$
		\item $ab-a'b' = ab-a'b+a'b-a'b' = (a-a')b + a'(b-b')$
	\end{enumerate}
	Thus property 1 and 2 are true.
\end{proof}
\begin{proposition}
	Let $n\in\mb Z$.
	\begin{enumerate}
		\item $[a]+[b] = [b] + [a]$ and $[a][b] = [b][a]$.
		\item $([a]+[b])+[c] = [a] + ([b]+[c])$ and $([a][b])[c] = [a]([b][c])$.
		\item $[a]+[0] = [a]$ and $[a]\cdot [1] =[a]$.
		\item $[a]([b]+[c]) = [a][b] + [a][c]$.
		\item $[a]+[-a] = [0]$.
		\item Let $a\neq 0$. Then $[a]$ has a multiplicative inverse in $\mb Z_n \Llr \gcd(a,n) = 1$.
		\begin{proof}
			\ul{$\La$}: Assume $\gcd(a,n) = 1$. Then $\exists r,s\in\mb Z$ such that $ar+ns = 1$ implies $ar+ns \equiv 1\pmod n$ or $ar\equiv 1\pmod n$. Thus $[a][r] = [ar] = [1]$ and $[a]\inv = [r]$.

			\ul{$\Ra$}: Assume $[a]$ has a multiplicative inverse in $\mb Z_n$: $\exists r\in\mb Z$ such that $[a][r] = [1]$ or $[ar] = [1]$, implying $ar\equiv 1\pmod n$. Thus $ar-1 = ns$ for some $s\in\mb Z$. Rewrite as $ar+n(-s) = 1 \implies \gcd(a,n) = 1$.
		\end{proof}
	\end{enumerate}
\end{proposition}

\subsection{Symmetries}

\begin{example}
	[Triangle] Let $\Delta ABC$ be a triangle.
	\begin{itemize}
		\item $id$ is identity symmetry, leave things as they are.
		\item $\rho_1$ is rotation by one vertex.
		\item $\rho_2$ is $\rho_1$ applied twice.
		\item $\mu_1$ is a reflection about $A$'s axis.
		\item $\mu_2$ is a reflection about $B$'s axis.
		\item $\mu_3$ is a reflection about $C$'s axis.
	\end{itemize}
	If you apply a sequence of symmetries, the result must also be one of these.
	Symmetry: function $\{A,B,C\}\ra\{A,B,C\}$.
	$\mu_1(A) = A$, $\mu_2(B) = B$, $\mu_3(C) = C$.
	Compose symmetries $\llr$ composition of functions $\rho_1\circ \mu _1$.

	Multiplication table (Cayley table):
	$$\begin{array}{c|cccccc}
		\circ & id & \rho_1 & \rho_2 & \mu_1 & \mu_2 & \mu_3 \\ 
		\hline
		id & id & \rho_1 & \rho_2 & \mu_1 & \mu_2 & \mu_3 \\ 
		\rho_1 & \rho_1 & \rho_2 & id & \mu_3 & \mu_1 & \mu_2 \\ 
		\rho_2 & \rho_2 & id & \rho_1 & \mu_2 & \mu_3 & \mu_1 \\ 
		\mu_1 & \mu_1 & \mu_2 & \mu_3 & id & \rho_1 & \rho_2 \\ 
		\mu_2 & \mu_2 & \mu_3 & \mu_1 & \rho_2 & id & \rho_1 \\ 
		\mu_3 & \mu_3 & \mu_1 & \mu_2 & \rho_1 & \rho_2 & id
	\end{array}$$
	All of the elements can be generated by two elements, one of each $\rho_1,\mu_1$.
	\begin{align*}
		id &= \rho_1\circ \rho_1 \circ \rho_1 = \rho_1^3 \\
		   &= \mu_1\circ\mu_1 = \mu_1^2 \\
		\rho_2 &= \rho_1\circ\rho_1 = \rho_1^2 \\
		\mu_2 &= \mu_1\circ\rho_1 = \mu_1\rho_1 \\
		\mu_3 &= \mu_1\rho_1^2
	\end{align*}
	The table can be rewritten as:
	$$\begin{array}{c|cccccc}
		\circ & id & \rho_1 & \rho_1^2 & \mu_1 & \mu_1\rho_1 & \mu_1\rho_1^2 \\
     	\hline
     	id & id & \rho_1 & \rho_2 & \mu_1 & \mu_2 & \mu_3 \\
     	\rho_1 & \rho_1 & \rho_2 & id & \mu_3 & \mu_1 & \mu_2 \\
	 	\rho_1^2 & \rho_2 & id & \rho_1 & \mu_2 & \mu_3 & \mu_1 \\
     	\mu_1 & \mu_1 & \mu_2 & \mu_3 & id & \rho_1 & \rho_2 \\
     	\mu_1\rho_1 & \mu_2 & \mu_3 & \mu_1 & \rho_2 & id & \rho_1 \\
     	\mu_1\rho_1^2 & \mu_3 & \mu_1 & \mu_2 & \rho_1 & \rho_2 & id
 	\end{array}$$
	Generators: $\rho_1,\mu_1$.
	Relations: $\rho_1^3 id$, $\mu_1^2 = id$, $\rho_1\mu_1 = \mu_1\rho_1^2$.

	Define the gorup by generators and relations: $D_6 = <\rho_1,\mu_1 \mid \rho_1^2 = id,\ \mu_1^2 = id,\ \rho_1\mu_1 = \mu_1\rho_1^2 >$.

	So $\mu_1\rho_1\mu_1\rho_1^2 = \mu_1\mu_1\rho_1^2\rho_1^2 = \mu_1^2\rho_1^4 = \rho_1$.
\end{example}

\subsection{Binary operations}

\begin{definition}
	A binary operation $\circ$ on a set $S$ is a mapping $\circ:S\times S\ra S$.
	\begin{enumerate}
		\item $(a,b)$ and $(b,a)$ are different ordered pairs (if $a\neq b$), so $a\circ b$ and $b\circ a$ could be different. $$\circ(a,b) = a\circ b.$$
		\item $a\circ b$ exists for every $a,b\in S$.
		\item $a\circ b$ is well-defined.
		\item $a\circ b\in S$ for every $a,b\in S$, so $S$ is \ul{closed} under the operation.
	\end{enumerate}
\end{definition}
\begin{example}
	$S = \mb R$, $a\circ b = a/b$. Not a binary operation, since $(a,0)\ra a\circ 0$ does not exist.
\end{example}
\begin{example}
	$S = \mb R$, $a\circ b = $ number whose square is $|ab|$. Take $(2,8) \ra 2\circ 8 = 4,-4$, so not a binary op.
\end{example}
\begin{example}
	$S = \mb N$, $a\circ b = a-b$. Not a binary operation since $(1,2)\ra 1\circ 2=1-2 = -1\notin S$.
\end{example}
\begin{example}
	$S = \mb N$, $a\circ b = a+b$, $\checkmark$. It is a binary operation since it is closed.
\end{example}
\begin{example}
	$X$ any set, $S = \mc P(x) = \{A\mid A\subseteq X\}$. Consider $X = \{\Box,\Delta\}$, then $\mc P(x) = \{\emptyset,\{\Box\},\{\Delta\},X\}$.
\end{example}
\begin{example}
	$A\circ B = A\cup B$, $\checkmark$. Since $A,B\subseteq X \implies A\cup B \subseteq X$.

	$A*B = A\cap B$, $\checkmark$. Since $A,B\subseteq X \implies A\cap B\subseteq X$.
\end{example}

\begin{definition}
	A binary operation $\circ$ on a set $S$ is
	\begin{enumerate}
		\item \ul{commutative} if $a\circ b = b\circ a$ for all $a,b\in S$.
		\item \ul{associative} if $(a\circ b)\circ c = a\circ(b\circ c)$ for all $a,b,c\in S$.
	\end{enumerate}
\end{definition}
\begin{example}
	$S = \mb N$, $a\circ b = \min\{a,b\}$. Since $a\circ b = \min\{a,b\}$ and $b\circ a = \min\{b,a\}$, commutative $\checkmark$.
	Since $(a\circ b)\circ c = \min\{a,b\}\circ c = \min\{\min\{a,b\},c\}$ and $a\circ(b\circ c) = a\circ\min\{b,c\} = \min\{a,\min\{b,c\}\}$ are both equal to $\min\{a,b,c\}$, associative $\checkmark$.
\end{example}
\begin{example}
	$S = \mb N$, $a\circ b = a$. Since $1\circ 2 = 1$ but $2\circ 1 = 2$, not commutative.
	$(a\circ b)\circ c = a\circ c = a$ and $a\circ(b\circ c) = a\circ b = a$, associative $\checkmark$.
\end{example}
\begin{example}
	$S = \mb Z$, $a\circ b = a-b$. Since $1\circ 2 = 1-2 = -1$ but $2\circ 1 = 2 = 2-1 = 1$, not commutative.
	Since $(a\circ b)\circ c = (a-b)\circ c = (a-b)-c = a-b-c$ and $a\circ(b\circ c) = a\circ(b-c) = a-(b-c) = a-b+c$, it is not associative because $(0\circ 0)\circ 1 = -1$ and $0\circ(0\circ 1) = 1$.
\end{example}
\begin{example}
	$S = \mb Z$, $a\circ b = 3(a+b)$. Then $a\circ b = 3(a+b)$ and $b\circ a = 3(b+a)$, commutative $\checkmark$.
	Also, $(a\circ b)\circ c = [3(a+b)]\circ c = 3(3(a+b)+c) = 9a+9b + 3c$ and $a\circ(b\circ c) = a\circ[3(b+c)] = 3(a+3(b+c)) = 3a+9b+9c$, not associative.
\end{example}
\begin{example}
	$S = M_n(\mb R)$ $n\times n$ matrices. $A\circ B = AB$. Then $AB\neq BA$, not commutative in general. However, $(AB)C = A(BC)$, associative $\checkmark$.
\end{example}

\begin{example}
	[HW Review] $ar+bs = 1 \implies \gcd(a,b) = 1$ since $d\mid a$ and $d\mid b$ and $d\mid 1 \implies \gcd(a,b) = 1$. However, $3\cdot 1 + 2\cdot(-1) = 1 \implies 3\cdot d + 2\cdot(-d) = d$.
\end{example}

\subsection{Grous: definition and examples}

\begin{definition}
	A nonempty set $G$ with a binary operation $\circ$ is a \ul{group} if
	\begin{enumerate}
		\item $\circ$ is associative
		\item there exists an element $e\in G$ such that $a\circ e = e\circ a = a$ for all $a\in G$ (identity element)
		\item for every $a\in G$ there exists $b\in G$ such that $a\circ b = b\circ a = e$ (inverse of $a$)
	\end{enumerate}
	\ul{Notation}: $(G,\circ)$ or $G$.
\end{definition}
\begin{definition}
	A group $(G,\circ)$ is \ul{abelian} if $\circ$ is commutative.
\end{definition}
\begin{remark}
	An operation can be commutative, a group can be abelian.
\end{remark}
\begin{example}
	$(\mb Z,+)$: associative $\checkmark$. $e=0 \checkmark$. inverse of $n$ is $-n$, $\checkmark$. commutative $\checkmark$. \ul{abelian group}.
	$(\mb Q, +)$, abelian group. So is $(\mb R,+)$, $(\mb C,+)$. For $(\mb R,\cdot)$ it is associative $\checkmark$, $e=1$, inverse of $x$ is $\frac{1}{x}$, no inverse for 0, \ul{not a group}. However, $(\mb R\backslash \{0\},\cdot)$ is an abelian group. $\mb R^\times$ is the multiplicative group of a field.
\end{example}
\begin{example}
	$(\mb Z_n,+)$: $[a]+[b] = [a+b]$ associative $\checkmark$. $e=[0]$. inverse of $[a]$ is $[-a]$. \ul{abelian group}.
\end{example}
\begin{example}
	$G = \{e\}$ is the trivial group, $e\circ e = e$, $e\inv = e$.
\end{example}
\begin{example}
	Symmetries of a regular $n$-gon. operation: composition. \ul{Claim}: non-abelian group. Notation can be $D_n$ or $D_{2n}$, they are called dihedral group.
\end{example}
\begin{example}
	$(M_n(\mb R),+)$, associative $\checkmark$, $e=0$ zero matrix, inverse of $A$ is $-A$, abelian group.
\end{example}
\begin{example}
	$(M_n(\mb R),\cdot)$: associative $\checkmark$, $e= I_n$, inverse of $A$ is $A\inv$ if $A$ is nonsingular. \ul{not a group}.
\end{example}
\begin{example}
	$GL_n(\mb R) = \{A\in M_n(\mb R) \mid \det A \neq 0\}$ nonsingular matrices. operation: multiplication. nonsingular $\cdot$ nonsingular $=$ nonsingular. associative $\checkmark$, $e = I_n$, inverse of $A$ is $A\inv$. \ul{nonabelian group} called \ul{general linear group}.
\end{example}
\begin{example}
	$(\mb Q^+,*)$, where $a*b = \frac{ab}{2}$.
	\begin{itemize}
		\item associative: $(a*b)*c = \frac{ab}{2} * c = \frac{ab/2\cdot c}{2} = \frac{abc}{4}$ and $a*(b*c) = a*\frac{bc}{2} = \frac{a\cdot bc/2}{2} = \frac{abc}{4}$.
		\item identity: $a*e = e*a = a$ or $\frac{ae}{2} = \frac{ea}{2} = a \implies e = 2$.
		\item inverses: $a*b = b*a = e$ or $\frac{ab}{2} = \frac{ba}{2} = 2 \implies b = \frac{4}{a}$ is inverse of $a$.
	\end{itemize}
	This is a \ul{group}.
\end{example}

\begin{definition}
	The \ul{order} of a group $(G,\circ)$ is the size of the set $G$: $|G|$. There are finite groups with $|G|<\infty$ and infinite groups with $|G|=\infty$.
\end{definition}

\end{document}
